---
title: "data"
output: html_document
date: '2022-05-14'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(scales)
library(glue)
library(plotly)
library(ggplot2)
library(stringr)
library(lubridate )
library(dplyr)
library(zoo)
library(hrbrthemes)
library(viridis)
library(leaflet)
library(rgdal)
library(GADMTools)
library(sf)
```

```{r}
b <- read.csv("data/bt.csv")
hls <- read_csv("data/hls.csv")
p <- read_csv("data/p.csv")
rls <- read_csv("data/rls.csv")
spp <- read_csv("data/spp.csv")
jenjang <- read_csv("data/jj.csv")
penyelesaian_daerah <- read_csv("data/penyelesaian-daerah.csv")
penyelesaian_kelamin <- read_csv("data/penyelesaian-kelamin.csv")
pt_kelamin <- read_csv("data/PT-kelamin.csv")
pt_pend <- read_csv("data/PT-pend.csv")
pt_bulan <- read_csv("data/PT-bln.csv")
pt_lulus <- read_csv("data/PT-lulus.csv")
```

```{r}
p
str(hls)
head(p)
head(rls)
head(spp)
```

```{r}
b <- b %>% mutate(nama_variabel = as.factor(nama_variabel),
         jenis_kelamin = as.factor(jenis_kelamin),
         provinsi = as.factor(provinsi) ,
         persen  = as.numeric(persen),
         tahun = as.factor(tahun))
hls <- hls %>% mutate(nama_variabel = as.factor(nama_variabel),
                    jenis_kelamin = as.factor(jenis_kelamin),
                    provinsi = as.factor(provinsi) ,
                    angka  = as.numeric(angka),
                    tahun = as.factor(tahun))
p <- p %>% mutate(nama_variabel = as.factor(nama_variabel),
         jenis_kelamin = as.factor(jenis_kelamin),
         provinsi = as.factor(provinsi) ,
         total  = as.numeric(total),
         tahun = as.factor(tahun))
rls <- rls %>%mutate(nama_variabel = as.factor(nama_variabel),
         jenis_kelamin = as.factor(jenis_kelamin),
         provinsi = as.factor(provinsi) ,
         angka  = as.numeric(angka),
         tahun = as.factor(tahun))
spp <- spp %>% mutate(nama_variabel = as.factor(nama_variabel),
         provinsi = as.factor(provinsi) ,
         persen  = as.numeric(persen),
         tahun = as.factor(tahun))
```

```{r}
str(p)
str(b)
str(hls)
str(rls)
str(spp)
```

```{r}
rls <- distinct(rls, nama_variabel,jenis_kelamin,tahun,angka,provinsi, .keep_all= TRUE)
rls
```

```{r}
hls_rls <- bind_rows(hls, rls)
hls_rls
```

```{r}
levels(hls$jenis_kelamin)
```


```{r}
rls_clean <- rls[rls$jenis_kelamin != "Laki-laki", ]
hls_clean <- hls[hls$jenis_kelamin != "Laki-laki", ]
```

```{r}
hls_rls_clean <- bind_rows(hls_clean, rls_clean)
hls_rls_clean <- hls_rls_clean %>% mutate(label=glue("Lama : {angka}"))
hls_rls_clean
```


```{r}
plot1 <- ggplot(hls_rls_clean, fill = nama_variabel) + 
  geom_point(aes(x = tahun, y = angka, fill = factor(nama_variabel, levels=c("Rata-rata Lama Sekolah","Angka Harapan Lama Sekolah")),text = label),
           stat = "identity", position = "dodge", show.legend = F) +
  scale_fill_viridis(discrete = T, option = "C") +
  scale_y_continuous("Count", breaks = seq(5,35,5), limits=c(0,20),
                     expand = expansion(mult = c(0.001, 0.001)),
                     sec.axis = dup_axis(name = NULL, labels = NULL)) +
  scale_x_discrete(expand = expansion(mult = c(0.0001, 0.0001)))+
  labs(title = NULL,
           x = NULL,
           y = NULL,
           fill = NULL) +
  theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1),
            plot.title = element_text(face = "bold"),
            panel.background = element_rect(fill = "#ffffff"),
            axis.line.y = element_line(colour = "grey"),
            axis.line.x = element_blank(),
            panel.grid = element_blank()) +
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position = "none")
plot1
```


 output$plothlsrls <- renderPlotly({
      hls_rls_clean <- hls_rls_clean  %>%  filter(tahun == input$input_tahun) %>% 
        mutate(label =glue("Lama(Tahun) : {angka}"))
    
      plothlsrls <- ggplot(hls_rls_clean) + 
        geom_bar(data =hls_rls_clean[hls_rls_clean$nama_variabel == "Angka Harapan Lama Sekolah",], aes(x = provinsi, y = angka, fill=provinsi, text = label), 
                 stat = "identity", position = "dodge", show.legend = F) +
        geom_line(data= hls_rls_clean[hls_rls_clean$nama_variabel == "Rata-rata Lama Sekolah",],aes(x = provinsi, y = angka ,group = 1)) +
        geom_point(data= hls_rls_clean[hls_rls_clean$nama_variabel == "Rata-rata Lama Sekolah",],aes(x = provinsi, y= angka)) +
        scale_fill_viridis(discrete = T, option = "C") +
        scale_y_continuous(breaks = seq(5,35,5), limits=c(0,20),
                           expand = expansion(mult = c(0.001, 0.001)),
                           sec.axis = dup_axis(name = NULL, labels = NULL)) +
        scale_x_discrete(expand = expansion(mult = c(0.0001, 0.0001)))+
        labs(title = NULL,
             x = NULL,
             y = NULL,
             fill = NULL) +
        theme(legend.title = element_blank(),
              axis.text.x = element_text(angle = 30, hjust = 1),
              plot.title = element_text(face = "bold"),
              panel.background = element_rect(fill = "#ffffff"),
              axis.line.y = element_line(colour = "grey"),
              axis.line.x = element_blank(),
              panel.grid = element_blank()) +
        theme(axis.text.x = element_text(angle = 45, hjust=1))
    
    ggplotly(plothlsrls, tooltip="text")


+
  geom_line(data= hls_rls_clean[hls_rls_clean$nama_variabel == "Rata-rata Lama Sekolah",],aes(x = tahun, y = angka ,group = 1, text = label)) +
  geom_point(data= hls_rls_clean[hls_rls_clean$nama_variabel == "Rata-rata Lama Sekolah",],aes(x = tahun, y= angka)) +

```{r}
ggplotly(plot1, tooltip="text")
```

```{r}
plothlsrls <- ggplot(hls_rls_clean, aes(x = tahun, y = angka, group=nama_variabel, color=nama_variabel, text = label)) + 
      geom_line() +
      scale_fill_viridis(discrete = T, option = "C") +
      labs(title = NULL,
           x = NULL,
           y = NULL,
           fill = NULL) +
      theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1),
            plot.title = element_text(face = "bold"),
            panel.background = element_rect(fill = "#ffffff"),
            axis.line.y = element_line(colour = "grey"),
            axis.line.x = element_blank(),
            panel.grid = element_blank()) +
      theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position = "top")
    
ggplotly(plothlsrls, tooltip="text")
```

plothlsrls <- ggplot(hls_rls_clean) + 
      geom_line(aes(x = tahun, y = angka, fill=nama_variabel, text = label), 
               stat = "identity", position = "dodge", show.legend = F) +
      facet_wrap(~jenis_kelamin) +
      scale_fill_viridis(discrete = T, option = "C") +
      scale_y_continuous(breaks = seq(5,35,5), limits=c(0,20),
                         expand = expansion(mult = c(0.001, 0.001)),
                         sec.axis = dup_axis(name = NULL, labels = NULL)) +
      scale_x_discrete(expand = expansion(mult = c(0.0001, 0.0001)))+
      labs(title = NULL,
           x = NULL,
           y = NULL,
           fill = NULL) +
      theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1),
            plot.title = element_text(face = "bold"),
            panel.background = element_rect(fill = "#ffffff"),
            axis.line.y = element_line(colour = "grey"),
            axis.line.x = element_blank(),
            panel.grid = element_blank()) +
      theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position = "top")
    
    ggplotly(plothlsrls, tooltip="text")


output$plothlsrls <- renderPlotly({
    hls_rls_clean <- hls_rls_clean %>% mutate(label =glue(" Kategori : {nama_variabel} 
                                                      Lama(Tahun) : {angka}")) %>%  filter(provinsi == input$input_provinsi) 
    
    plothlsrls <- ggplot(hls_rls_clean) + 
      geom_bar(aes(x = tahun, y = angka, fill=nama_variabel, text = label), 
                stat = "identity", position = "dodge", show.legend = F) +
      scale_fill_viridis(discrete = T, option = "C") +
      scale_y_continuous(breaks = seq(5,35,5), limits=c(0,20),
                         expand = expansion(mult = c(0.001, 0.001)),
                         sec.axis = dup_axis(name = NULL, labels = NULL)) +
      scale_x_discrete(expand = expansion(mult = c(0.0001, 0.0001)))+
      labs(title = NULL,
           x = NULL,
           y = NULL,
           fill = NULL) +
      theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1),
            plot.title = element_text(face = "bold"),
            panel.background = element_rect(fill = "#ffffff"),
            axis.line.y = element_line(colour = "grey"),
            axis.line.x = element_blank(),
            panel.grid = element_blank()) +
      theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position = "top")
    
    ggplotly(plothlsrls, tooltip="text")


```{r}
head(penyelesaian_daerah)
head(penyelesaian_kelamin)
head(b)
```

```{r}
penyelesaian_kelamin <- penyelesaian_kelamin %>%mutate(tingkat_pendidikan = as.factor(tingkat_pendidikan),
         jenis_kelamin = as.factor(jenis_kelamin),
         nama_variabel = as.factor(nama_variabel) ,
         persen  = as.numeric(persen),
         tahun = as.factor(tahun))
penyelesaian_daerah <- penyelesaian_daerah %>%mutate(tingkat_pendidikan = as.factor(tingkat_pendidikan),
         provonsi = as.factor(provonsi),
         nama_variabel = as.factor(nama_variabel) ,
         persen  = as.numeric(persen),
         tahun = as.factor(tahun))
```


```{r}

pdk <- bind_rows(penyelesaian_daerah, penyelesaian_kelamin)
pdk <- pdk %>% rename(provinsi = provonsi)
pdk
```


```{r}
pdk <- pivot_longer(data = pdk,
                           cols = c("tingkat_pendidikan","jenis_kelamin"),
                           names_to = "variabel")
pdk
```

```{r}
pdk <- pdk %>%
  rename(jenis = value)
```

```{r}
pdk <- pdk %>% na.omit()
pdk <- pdk %>%  mutate(label =glue("Variabel :{variabel}
                                   Sub Variabel : {jenis}
                                    Total % : {persen}"))
```

```{r}
pdk <- pdk  %>% select (-c("variabel", "nama_variabel"))
pdk 
``` 


```{r}
plotpdk <- pdk %>% ggplot(aes(x= provinsi, y= persen, fill = jenis)) +
      geom_line(aes(group=1),position = "identity", colour="black", size=0.2, alpha=.6) +
      geom_point(aes(text = label)) +
      scale_fill_viridis(discrete = TRUE, option = "C") +
      theme(legend.position="top")
plotpdk
```

```{r}
ggplotly(plotpdk, tooltip = "text")
```


```{r}
  ggplot(pdk, aes(x=provinsi, y= persen, fill=jenis)) +
    geom_area(position = "identity", colour="black", size=0.2, alpha=.4) +
    geom_point(aes(text=label)) +
    scale_fill_viridis(discrete = TRUE, option = "C") +
    ggtitle("Korban menurut Jenis Kelamin Per Bulan") +
    theme_minimal() +
    theme(legend.position="top")
```


```{r}
penyelesaian_daerah <- penyelesaian_daerah  %>% 
  mutate(label =glue(" Tingkat Pendidikan : {tingkat_pendidikan}
                       Total % : {persen}"))
```



```{r}
plotpd <- penyelesaian_daerah %>% ggplot(aes(x=tahun, y= persen, fill = tingkat_pendidikan)) +
    geom_area(position = "identity", colour="black", size=0.2, alpha=.4) +
    geom_point(aes(text = label)) +
    scale_fill_viridis(discrete = TRUE, option = "C") +
    ggtitle("Tingkat Penyelesaian di Masing-Masing Provinsi") +
    theme_minimal() +
    theme(legend.position="top")
plotpd
```

```{r}
pdk %>%
  filter(jenis %in% c("SD / Sederajat", "SMP / Sederajat", "SMA / Sederajat")) %>%
  ggplot(aes(x= provinsi, y=persen, color=jenis, fill=jenis)) +
    geom_density(alpha=0.6) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none"
    ) +
    ylab("") +
    xlab("Assigned Probability (%)")
```

```{r}
b
head(pt_kelamin)
head(pt_pend)
pt_bulan
pt_lulus
```

```{r}
pt_kelamin <- pt_kelamin %>% select(-c("nama_variabel")) %>% mutate(
         jenis_kelamin = as.factor(jenis_kelamin),
         persen  = as.numeric(persen),
         tahun = as.factor(tahun))
pt_pend <- pt_pend %>% select(-c("nama_variabel")) %>% mutate(
         tingkat_pendidikan = as.factor(tingkat_pendidikan),
         persen  = as.numeric(persen),
         tahun = as.factor(tahun))
pt_bulan <- pt_bulan %>% filter_all(any_vars(. %in% c('Jumlah (Ribu orang)'))) %>% 
  select(-c("nama_variabel", "satuan")) %>%  mutate(
         persen  = as.numeric(persen),
         tahun = as.factor(tahun),
         bulan = as.factor(bulan))
pt_lulus <- pt_lulus %>% select(-c("nama_variabel")) %>% mutate(
         tingkat_pendidikan = as.factor(tingkat_pendidikan),
         jumlah  = as.numeric(jumlah),
         tahun = as.factor(tahun),
         bulan = as.factor(bulan))

```

```{r}
pt_bulan <- pt_bulan %>%
  group_by(bulan) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = bulan, values_from = persen) %>%
  select(-row) %>% 
  rowwise() %>% mutate(total = mean(c(Februari, Agustus), na.rm=T)) %>%
  select(-c(Februari, Agustus))
pt_bulan <- distinct(pt_bulan, tahun,total, .keep_all= TRUE)
pt_bulan
```


```{r}
b_provinsi <- b[b$provinsi != "INDONESIA", ]
b_clean <- b_provinsi[b_provinsi$jenis_kelamin != "Laki-Laki", ]
b_clean$persen<-format(round(b_clean$persen,2),nsmall=2)
b_clean <- b_clean %>% mutate(persen = as.numeric(persen))
b_clean <- b_clean %>% select(-c("nama_variabel", "jenis_kelamin")) %>% mutate(label=glue(" Provinsi: {provinsi}
                                                                                          Total % : {persen}"))
b_clean
```

```{r}
b
```


```{r}
plotbclean <- b_clean %>% ggplot(aes(x= tahun, y=reorder(persen, tahun), group=1)) +
      geom_area(fill = "light blue", alpha=.3) +
      geom_line(color= "blue") +
      geom_point(aes(text = label)) +
      scale_fill_viridis(discrete = TRUE, option = "C") +
      labs(title = NULL,
           x = NULL,
           y = NULL,
           fill = NULL) +
      theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1),
            plot.title = element_text(face = "bold"),
            panel.background = element_rect(fill = "#ffffff"),
            axis.line.y = element_line(colour = "grey"),
            axis.line.x = element_blank(),
            panel.grid = element_blank()) +
      theme(legend.position="none", axis.text.x = element_text(angle = 45, hjust=1))

plotbclean
```

```{r}
ptk_ptb <- bind_rows(pt_kelamin, pt_bulan)
ptk_ptb <- pivot_wider(data = ptk_ptb, 
                       names_from = jenis_kelamin, 
                       values_from = c(persen, total))
ptk_ptb <- ptk_ptb %>% rename(persen_l = `persen_Laki - Laki`) %>%
  select(-c("total_Laki - Laki", "total_Perempuan", "persen_NA"))
ptk_ptb$persen_l <- ptk_ptb$persen_l / 100
ptk_ptb$persen_Perempuan <- ptk_ptb$persen_Perempuan / 100
ptk_ptb$perempuan <- ptk_ptb$total_NA * ptk_ptb$persen_Perempuan
ptk_ptb$perempuan <- format(round(ptk_ptb$perempuan,2),nsmall=2)
ptk_ptb$total_keseluruhan <- format(round(ptk_ptb$total_NA,2),nsmall=2)
ptk_ptb <- ptk_ptb %>% select(-c("persen_l", "persen_Perempuan", "total_NA"))
ptk_ptb <- pivot_longer(data = ptk_ptb,
                           cols = c("perempuan","total_keseluruhan"),
                           names_to = "variabel")
ptk_ptb$variabel <- as.factor(ptk_ptb$variabel)
ptk_ptb
```



```{r}
      plotptk_ptb <- ggplot(ptk_ptb, mapping = aes(x = tahun, y = reorder(value, tahun), group=variabel, color=variabel )) +
        geom_line(stat = "identity", show.legend = F) +
        geom_point() +
        scale_fill_viridis(discrete = T, option = "C") +
        labs(title = NULL,
             x = NULL,
             y = NULL,
             fill = NULL) +
        theme(legend.title = element_blank(),
              axis.text.x = element_text(angle = 30, hjust = 1),
              plot.title = element_text(face = "bold"),
              panel.background = element_rect(fill = "#ffffff"),
              axis.line.y = element_line(colour = "grey"),
              axis.line.x = element_blank(),
              panel.grid = element_blank()) +
        theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust=1), )
plotptk_ptb
```

```{r}
levels(pt_pend$tingkat_pendidikan)
pt_pend
```


```{r}
plotptp <- ggplot(pt_pend, mapping = aes(x = tahun, y = reorder(persen, tahun), group=tingkat_pendidikan, color=tingkat_pendidikan )) +
        geom_line(stat = "identity", show.legend = F) +
        geom_point() +
        scale_fill_viridis(discrete = T, option = "C") +
        labs(title = NULL,
             x = NULL,
             y = NULL,
             fill = NULL) +
        theme(legend.title = element_blank(),
              axis.text.x = element_text(angle = 30, hjust = 1),
              plot.title = element_text(face = "bold"),
              panel.background = element_rect(fill = "#ffffff"),
              axis.line.y = element_line(colour = "grey"),
              axis.line.x = element_blank(),
              panel.grid = element_blank()) +
        theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust=1), )
plotptp
```

```{r}
spp
head(p)
```

```{r}
p <- p[p$jenis_kelamin != "Laki-laki", ]
p <- p %>% select(-c("jenis_kelamin"))
p <- distinct(p, nama_variabel,tahun,total,provinsi, .keep_all= TRUE)
p_provinsi <- p[p$provinsi != "INDONESIA", ]
p_provinsi
```

```{r}
plotp <- p_spp_prov[p_spp_prov$nama_variabel == "Pengeluaran per Kapita",] %>% ggplot(aes(x= tahun, y=reorder(total, tahun), group=1)) +
      geom_col(fill = "blue", alpha=.3) +
      geom_line(color= "blue") +
      geom_point() +
      scale_fill_viridis(discrete = TRUE, option = "C") +
      labs(title = NULL,
           x = NULL,
           y = NULL,
           fill = NULL) +
      theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1),
            plot.title = element_text(face = "bold"),
            panel.background = element_rect(fill = "#ffffff"),
            axis.line.y = element_line(colour = "grey"),
            axis.line.x = element_blank(),
            panel.grid = element_blank()) +
      theme(legend.position="none", axis.text.x = element_text(angle = 45, hjust=1))
plotp
```

```{r}
spp <- distinct(spp, tahun,persen,provinsi, .keep_all= TRUE)
spp$persen<-format(round(spp$persen,2),nsmall=2)
spp_provinsi <- spp[spp$provinsi != "INDONESIA", ]
spp_provinsi
```
```{r}
p_spp_prov$persen <- as.numeric(p_spp_prov$persen)
p_spp_prov_a <- pivot_longer(data = p_spp_prov,
                           cols = c("total", "persen"),
                           names_to = "jumlah")
p_spp_prov_a  <- p_spp_prov_a %>% na.omit()
p_spp_prov_a <- p_spp_prov_a %>% select(-c(jumlah))
p_spp_prov_a <- pivot_wider(data = p_spp_prov_a, 
                       names_from = nama_variabel, 
                       values_from = c(value))
p_spp_prov_a <- distinct(p_spp_prov_a, tahun,`Pengeluaran per Kapita`,`Sumbangan Pendapatan Perempuan`,provinsi, .keep_all= TRUE)
p_spp_prov_a <- p_spp_prov_a %>% rename( pengeluaran_per_kapita =`Pengeluaran per Kapita`, sumbangan_pendapatan = `Sumbangan Pendapatan Perempuan`)
```

```{r}
pt_pend_2021 <- pt_pend %>% filter(tahun == "2021")
pt_pend_2021 <- pt_pend_2021 %>% mutate(label = glue("Tingkat Pengangguran Terbuka dalam Persen : {persen}"))
```


```{r}
plot2 <- ggplot(pt_pend_2021, mapping = aes(x = reorder(tingkat_pendidikan, persen), y = persen, fill = tingkat_pendidikan, group = 1, text = label)) +
      geom_bar(stat = "identity", show.legend = F) +
      scale_fill_viridis(discrete = T, option = "C") +
      scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
      labs(title = NULL,
           x = NULL,
           y = NULL,
           fill = NULL) +
      theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1),
            plot.title = element_text(face = "bold"),
            panel.background = element_rect(fill = "#ffffff"),
            axis.line.y = element_line(colour = "grey"),
            axis.line.x = element_blank(),
            panel.grid = element_blank()) +
      theme(legend.position = "none", axis.text.x = element_text(angle = 30, hjust=1), )
plot2
```


```{r}
scat_plot <- ggplot(p_spp_prov_a, aes(x = pengeluaran_per_kapita , y = sumbangan_pendapatan)) +
        geom_point(aes(color = tahun)) +
        scale_y_continuous(limits = c(0,18)) +
        scale_color_manual(values = c("#DF0101", "#04B486")) +
        labs(title = "HDI on Ecological Footprint",
             y = "Ecological Footprint",
             x = "Human Development Index") +
        theme(plot.title = element_text(face = "bold", size = 14, hjust = 0),
              panel.background = element_rect(fill = "#ffffff"), 
              panel.grid.major.x = element_line(colour = "grey"),
              panel.grid.major.y = element_line(colour = "grey"),
              axis.line.x = element_line(color = "grey"),
              axis.line.y = element_line(color = "grey"),
              axis.text = element_text(size = 10, colour = "black"),
              legend.title = element_blank())
scat_plot
```


```{r}
plot1 <- ggplot(p_spp_prov, fill = nama_variabel) + 
  geom_point(aes(x = tahun, y = angka, fill = factor(nama_variabel, levels=c("Rata-rata Lama Sekolah","Angka Harapan Lama Sekolah")),text = label),
           stat = "identity", position = "dodge", show.legend = F) +
  scale_fill_viridis(discrete = T, option = "C") +
  scale_y_continuous("Count", breaks = seq(5,35,5), limits=c(0,20),
                     expand = expansion(mult = c(0.001, 0.001)),
                     sec.axis = dup_axis(name = NULL, labels = NULL)) +
  scale_x_discrete(expand = expansion(mult = c(0.0001, 0.0001)))+
  labs(title = NULL,
           x = NULL,
           y = NULL,
           fill = NULL) +
  theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1),
            plot.title = element_text(face = "bold"),
            panel.background = element_rect(fill = "#ffffff"),
            axis.line.y = element_line(colour = "grey"),
            axis.line.x = element_blank(),
            panel.grid = element_blank()) +
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position = "none")
plot1
```

```{r}
p_spp <- bind_rows(p, spp)
p_spp_prov <- p_spp[p_spp$provinsi != "INDONESIA", ]
p_spp_prov[p_spp_prov$nama_variabel == "Pengeluaran per Kapita",]
p_spp_prov
```


```{r}
top5hlsrls <- pivot_wider(data = hls_rls_clean, 
                       names_from = nama_variabel, 
                       values_from = c(angka))
top5hlsrls <- top5hlsrls %>% rename(rata = `Rata-rata Lama Sekolah`, harapan = `Angka Harapan Lama Sekolah`)
top5hlsrls %>% arrange(-harapan, rata)
```
```{r}
hlsarr <- hls_clean %>% arrange(-angka) %>% filter(tahun == "2021") %>% rename(harapan = angka) %>% select(-c(nama_variabel, jenis_kelamin, tahun))
rlsarr <- rls_clean %>% arrange(-angka) %>% filter(tahun == "2021") %>% rename(realitas = angka)%>% select(-c(nama_variabel, jenis_kelamin, tahun))
hlsrlsarr <- bind_rows(hlsarr, rlsarr)
hlsrlsarr <- pivot_longer(data = hlsrlsarr,
                           cols = c("harapan","realitas"),
                           names_to = "variabel")
hlsrlsarr <- hlsrlsarr %>% na.omit()
hlsrlsarr$variabel <- as.factor(hlsrlsarr$variabel)
hlsrlsarr <- hlsrlsarr %>% mutate(label = glue("Provinsi : {provinsi}
                                      Variabel : {variabel}
                                      Angka (Tahun) : Value"))
rlsarr
```

```{r}
    
    plotprovinsi <- p_spp_prov %>% ggplot(mapping = aes(x= tahun, y=reorder(total, tahun), group=1)) +
      geom_col(fill = "blue", alpha=.3) +
      geom_line(color= "blue") +
      geom_point() +
      scale_fill_viridis(discrete = TRUE, option = "C") +
      labs(title = NULL,
           x = NULL,
           y = NULL,
           fill = NULL) +
      theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1),
            plot.title = element_text(face = "bold"),
            panel.background = element_rect(fill = "#ffffff"),
            axis.line.y = element_line(colour = "grey"),
            axis.line.x = element_blank(),
            panel.grid = element_blank()) +
      theme(legend.position="none", axis.text.x = element_text(angle = 45, hjust=1))
plotprovinsi
```



```{r}
    plothr <- hlsrlsarr %>% ggplot(mapping = aes(x= reorder(provinsi,value), y=value, fill = variabel, group= variabel, name = variabel)) +
      geom_area(position = "identity", colour="black", size=0.2, alpha=.4) +
      geom_point(aes(text = label)) +
      scale_fill_viridis(discrete = TRUE, option = "C") +
      labs(title = NULL,
           x = NULL,
           y = NULL,
           fill = NULL) +
      theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1),
            plot.title = element_text(face = "bold"),
            panel.background = element_rect(fill = "#ffffff"),
            axis.line.y = element_line(colour = "grey"),
            axis.line.x = element_blank(),
            panel.grid = element_blank()) +
      theme(legend.position="none", axis.text.x = element_text(angle = 45, hjust=1))
plothr
```
```{r}
hlsrlsarr %>%
  ggplot(aes(x=provinsi, y= value, fill=variabel)) +
    geom_area(position = "identity", colour="black", size=0.2, alpha=.4) +
    geom_point(aes(text=label)) +
    scale_fill_viridis(discrete = TRUE, option = "C") +
    ggtitle("Korban menurut Jenis Kelamin Per Bulan") +
    theme_minimal() +
    theme(legend.position="top")
```

```{r}
head(hlsrlsarr[hlsrlsarr$variabel == "harapan", ],1)
hlsrlsarr %>% filter(variabel == 'realitas') %>%
           filter(value == max(value, na.rm = TRUE)) %>% select(value)
```


## Including Plots

checkboxGroupInput (
                                inputId = "input_ptp",
                                label = "Pilih Variabel",
                                choices = c("Tidak/Belum Pernah Sekolah/Belum Tamat & Tamat SD" = "Tidak/Belum Pernah Sekolah/Belum Tamat & Tamat SD", 
                                            "SMP"="SMP", 
                                            "SMA umum"= "SMA umum", 
                                            "SMA Kejuruan"="SMA Kejuruan", 
                                            "	Diploma I/II/III"="	Diploma I/II/III", 
                                            "Universitas" = "Universitas"),
                                selected = "Tidak/belum pernah sekolah")


You can also embed plots, for example:

    if(identical(input$input_ptp, "Tidak/Belum Pernah Sekolah/Belum Tamat & Tamat SD")) {
    pt_pend <- pt_pend %>% filter(tingkat_pendidikan == "Tidak/Belum Pernah Sekolah/Belum Tamat & Tamat SD") %>% mutate(label = glue(" Tingkat Pendidikan : {tingkat_pendidikan}
                                                                                                                Total % : {persen}"))

    plotptp <- ggplot(pt_pend, mapping = aes(x = tahun, y = reorder(persen, tahun), group=tingkat_pendidikan, color=tingkat_pendidikan )) +
      geom_line(stat = "identity", show.legend = F) +
      geom_point(aes(text = label)) +
      scale_fill_viridis(discrete = T, option = "C") +
      labs(title = NULL,
           x = NULL,
           y = NULL,
           fill = NULL) +
      theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1),
            plot.title = element_text(face = "bold"),
            panel.background = element_rect(fill = "#ffffff"),
            axis.line.y = element_line(colour = "grey"),
            axis.line.x = element_blank(),
            panel.grid = element_blank()) +
      theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust=1), )
        } else { 
          if(identical(input$input_ptp, "SMP")) {
            pt_pend <- pt_pend %>% filter(tingkat_pendidikan == "SMP") %>% mutate(label = glue(" Tingkat Pendidikan : {tingkat_pendidikan}
                                                                                                  Total % : {persen}"))
            
            plotptp <- ggplot(pt_pend, mapping = aes(x = tahun, y = reorder(persen, tahun), group=tingkat_pendidikan, color=tingkat_pendidikan )) +
              geom_line(stat = "identity", show.legend = F) +
              geom_point(aes(text = label)) +
              scale_fill_viridis(discrete = T, option = "C") +
              labs(title = NULL,
                   x = NULL,
                   y = NULL,
                   fill = NULL) +
              theme(legend.title = element_blank(),
                    axis.text.x = element_text(angle = 30, hjust = 1),
                    plot.title = element_text(face = "bold"),
                    panel.background = element_rect(fill = "#ffffff"),
                    axis.line.y = element_line(colour = "grey"),
                    axis.line.x = element_blank(),
                    panel.grid = element_blank()) +
              theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust=1), )
          } else { 
            if(identical(input$input_ptp, "SMA umum")) {
              pt_pend <- pt_pend %>% filter(tingkat_pendidikan == "SMA umum") %>% mutate(label = glue("Tingkat Pendidikan : {tingkat_pendidikan}
                                                                                                            Total % : {persen}"))
              
              plotptp <- ggplot(pt_pend, mapping = aes(x = tahun, y = reorder(persen, tahun), group=tingkat_pendidikan, color=tingkat_pendidikan )) +
                geom_line(stat = "identity", show.legend = F) +
                geom_point(aes(text = label)) +
                scale_fill_viridis(discrete = T, option = "C") +
                labs(title = NULL,
                     x = NULL,
                     y = NULL,
                     fill = NULL) +
                theme(legend.title = element_blank(),
                      axis.text.x = element_text(angle = 30, hjust = 1),
                      plot.title = element_text(face = "bold"),
                      panel.background = element_rect(fill = "#ffffff"),
                      axis.line.y = element_line(colour = "grey"),
                      axis.line.x = element_blank(),
                      panel.grid = element_blank()) +
                theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust=1), )
            } else { 
              if(identical(input$input_ptp, "SMA Kejuruan")) {
                pt_pend <- pt_pend %>% filter(tingkat_pendidikan == "SMA Kejuruan") %>% mutate(label = glue("Tingkat Pendidikan : {tingkat_pendidikan}
                                                                                                                  Total % : {persen}"))
                
                plotptp <- ggplot(pt_pend, mapping = aes(x = tahun, y = reorder(persen, tahun), group=tingkat_pendidikan, color=tingkat_pendidikan )) +
                  geom_line(stat = "identity", show.legend = F) +
                  geom_point(aes(text = label)) +
                  scale_fill_viridis(discrete = T, option = "C") +
                  labs(title = NULL,
                       x = NULL,
                       y = NULL,
                       fill = NULL) +
                  theme(legend.title = element_blank(),
                        axis.text.x = element_text(angle = 30, hjust = 1),
                        plot.title = element_text(face = "bold"),
                        panel.background = element_rect(fill = "#ffffff"),
                        axis.line.y = element_line(colour = "grey"),
                        axis.line.x = element_blank(),
                        panel.grid = element_blank()) +
                  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust=1), )
                } else { 
                  if(identical(input$input_ptp, "	Diploma I/II/III")) {
                    pt_pend <- pt_pend %>% filter(tingkat_pendidikan == "	Diploma I/II/III") %>% mutate(label = glue("Tingkat Pendidikan : {tingkat_pendidikan}
                                                                                                                    Total % : {persen}"))
                    
                    plotptp <- ggplot(pt_pend, mapping = aes(x = tahun, y = reorder(persen, tahun), group=tingkat_pendidikan, color=tingkat_pendidikan )) +
                      geom_line(stat = "identity", show.legend = F) +
                      geom_point(aes(text = label)) +
                      scale_fill_viridis(discrete = T, option = "C") +
                      labs(title = NULL,
                           x = NULL,
                           y = NULL,
                           fill = NULL) +
                      theme(legend.title = element_blank(),
                            axis.text.x = element_text(angle = 30, hjust = 1),
                            plot.title = element_text(face = "bold"),
                            panel.background = element_rect(fill = "#ffffff"),
                            axis.line.y = element_line(colour = "grey"),
                            axis.line.x = element_blank(),
                            panel.grid = element_blank()) +
                      theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust=1), )
                } else { 
                  if(identical(input$input_ptp, "Universitas")) {
                    pt_pend <- pt_pend %>% filter(tingkat_pendidikan == "Universitas") %>% mutate(label = glue("Tingkat Pendidikan : {tingkat_pendidikan}
                                                                                                                Total % : {persen}"))
                    
                    plotptp <- ggplot(pt_pend, mapping = aes(x = tahun, y = reorder(persen, tahun), group=tingkat_pendidikan, color=tingkat_pendidikan )) +
                      geom_line(stat = "identity", show.legend = F) +
                      geom_point(aes(text = label)) +
                      scale_fill_viridis(discrete = T, option = "C") +
                      labs(title = NULL,
                           x = NULL,
                           y = NULL,
                           fill = NULL) +
                      theme(legend.title = element_blank(),
                            axis.text.x = element_text(angle = 30, hjust = 1),
                            plot.title = element_text(face = "bold"),
                            panel.background = element_rect(fill = "#ffffff"),
                            axis.line.y = element_line(colour = "grey"),
                            axis.line.x = element_blank(),
                            panel.grid = element_blank()) +
                      theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust=1), )
                  }}}}}}


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
